\Large \bf Cyclone: High Availability for Persistent Key Value Stores

### 2017-07-27 paper/paper.tex
Updates.

\title{\Large \bf Cyclone: Warp Speed Replication for Persistent Key Value Stores}

@@

The first key contribution of Cyclone is demonstrating the benefit of a small
amount of non-volatile memory when combined with the flash SSD based persistent
logs.  The primary cause of slowdown when turning on a write ahead log on a
single machine in Figure~\ref{fig:problem} is the need to do synchronous block
IO for every update. Mordern NVMe SSDs advertise a synchronous write latency of
the order of around 20us, capping throughput at 50K ops/sec, even with write coalescing 
in the SSDs own internal write cache. Cyclone avoids this by making use of directly attached persistent
memory on the server. This can be in the form of NVDIMMs (DIMMs with an added
ultracapacitor to dump state to a small amount of attached flash) or newer forms
of persistent memory~\cite{pmfs}. For current deployments, NVDIMMs are expensive
- retailing at approximately 2X the price of a standard DIMM [find ref]. This
means that they will likely be available in small quantities. We compensate for
this by periodically draining the NVDIMM log into a log placed on a standard SSD
on the system.

@@

\section{Storage}
\label{sec:storage}
Storage devices such as flash-based SSDs export a block IO interface. Appending
a small update to the write-ahead log requires a synchronous access to the SSD. Even with 
write coalescing using a write cache on the SSD, the round  trip time to a current generation NVMe 
SSD such as the one we use in this paper~\cite{ssd_spec}, is of the order of 20 us, limiting the throughput to under 50K ops/sec.

This leads to a high baseline latency at even low load, as shown in
Figure~\ref{fig:problem}. This problem leads many system designers to
underprovision key-value store shards ensuring that it operates at moderate to
high load, using group-commit to batch additions to the log. They therefore pay
a price in latency to ensure the bandwidth to storage is fully utilized to
prevent it from becoming a bottleneck.

@@

Cyclone takes a software only approach to improving replication performance
using commodity network hardware. In contrast, Mojim~\cite{mojim} pairs directly
attached persistent memory with RDMA to replicate persistent data structures. A
comparison is hard since Mojim is designed for the hypothetical scenario where
there is enough persistent memory to hold all persistent data, while Cyclone is
designed for the practical case where both the write-ahead log and key value
store data structures need to be maintained largely on flash. Nevertheless we
can compare their replication performance for the case where we use only the NVM
log in Cyclone. On 40Gbps infiniband, Mojim reported being able to mirror data
at 4GB/s. In contrast, Cyclone can replicate data three ways using 4*10 Gbps
commodity ethernet NICs at a peak of roughly 3.7 GB/s. Notably this is while
Cyclone is also running the RAFT consensus protocol, while Mojim is doing simple
mirroring of main memory. One can conclude that network offload via RDMA and
Infiniband is not an a-priori requirement for high throughput replication. This
is made possible by Cyclone's approach that relieves the CPU from the burden of
data copies. Of course, the latency of a hop on commodity ethernet at 5 us with
our DPDK stack is larger than the close to 2us latency with
Infiniband. Depending on the use case for the key value store application the
extra 6us with ethernet for the round trip to the replica might be a concern for
some applications. We note that mapping the commutativity of key value store
interfaces to multiple logs in Cyclone is a technique that can also be applied
to RDMA. The authors of Mojim confirm that concurrency is a significant
determinant of performance when replicating over RDMA.

Consensus in a box~\cite{consensus_box} implements the Zookeeper atomic
broadcast protocol together with a NIC on an FPGA. That work ignored durability,
focusing purely on network performance for replication. Latency numbers are
excellent due to the FPGA that cuts out the path from CPU caches to the NIC and
are reported to be as low as 3us for a round trip from leader to follower
replica compared to the 6us we observe with DPDK. Notably however they reported
7us when using TCP/IP from the FPGA rather than their own connection oriented
network protocol. On the other hand their peak replication throughput is 6M
replications/sec on 66 Gbps of aggregate network bandwith, comparable to the
same mark obtained by Cyclone on an aggregate 40 Gbps of network bandwidth when
not using the log on the SSD. We believe the general applicability afforded by
our software only solution together with the fact that we provide persistence by
design makes it a compelling alternative to an FPGA based solution even given
the larger latency.

>>>>

\title{\Large \bf Cyclone: High Availability for Persistent Key Value Stores}

@@

The first key contribution of Cyclone is demonstrating the benefit of a small
amount of non-volatile memory when combined with the flash SSD based persistent
logs.  The primary cause of slowdown when turning on a write ahead log on a
single machine in Figure~\ref{fig:problem} is the need to do synchronous block
IO for every update \emph{regardless of the size of the update}. SSDs provide
the maximum throughput when writing large chunks (usually more than 4KB) of
data. For small updates, the synchronous write latency dominates the time for
IO.  Cyclone avoids this by making use of directly attached persistent memory on
the server. This can be in the form of NVDIMMs (DIMMs with an added
ultracapacitor to dump state to a small amount of attached flash) or newer forms
of persistent memory~\cite{pmfs}. For current deployments, NVDIMMs are expensive
- retailing at approximately 2X the price of a standard DIMM [find ref]. This
means that they will likely be available in small quantities. We compensate for
this by periodically draining the NVDIMM log into a log placed on a standard SSD
on the system.

@@

\section{Storage}
\label{sec:storage}
Storage devices such as flash-based SSDs export a block IO interface. Appending
a small update to the write-ahead log requires a synchronous access to the
SSD. Even with write coalescing using a write cache on the SSD, the round trip
time to a current generation NVMe SSD such as the one we use in this
paper~\cite{ssd_spec}, is of the order of 20 us, limiting the throughput to
under 50K ops/sec. This leads to a high baseline latency at even low load, as
shown in Figure~\ref{fig:problem}. This problem leads many system designers to
underprovision key-value store shards ensuring that it operates at moderate to
high load, using group-commit to batch additions to the log. They therefore pay
a price in latency to ensure the bandwidth to storage is fully utilized to
prevent it from becoming a bottleneck.

@@

Cyclone takes a software only approach to improving replication performance
using commodity network hardware. In contrast, Mojim~\cite{mojim} pairs directly
attached persistent memory with RDMA to replicate persistent data structures. A
comparison is hard since Mojim is designed for the hypothetical scenario where
there is enough persistent memory to hold all persistent data, while Cyclone is
designed for the practical case where both the write-ahead log and key value
store data structures need to be maintained largely on flash. Nevertheless we
can compare their replication performance for the case where we use only the NVM
log in Cyclone. On 40Gbps infiniband, Mojim reported being able to mirror data
at 4GB/s. In contrast, Cyclone can replicate data three ways using 4*10 Gbps
commodity ethernet NICs at a peak of roughly 3.7 GB/s. This is with
Cyclone is also running the RAFT consensus protocol, while Mojim is doing simple
mirroring of main memory. One can conclude that network offload via RDMA and
Infiniband is not an a-priori requirement for high throughput replication. This
is made possible by Cyclone's approach that relieves the CPU from the burden of
data copies. Of course, the latency of a hop on commodity ethernet at 5 us with
our DPDK stack is larger than the close to 2us latency with
Infiniband. Depending on the use case for the key value store application the
extra 6us with ethernet for the round trip to the replica might be a concern for
some applications. We note that mapping the commutativity of key value store
interfaces to multiple logs in Cyclone is a technique that can also be applied
to RDMA. Mojim confirms that concurrency is a significant
determinant of performance when replicating over RDMA.

Consensus in a box~\cite{consensus_box} implements the Zookeeper atomic
broadcast protocol together with a NIC on an FPGA. That work ignored durability,
focusing purely on network performance for replication. Latency is
excellent due to the FPGA that cuts out the path from CPU caches to the NIC and
are reported to be as low as 3us for a round trip from leader to follower
replica compared to the 6us we observe with DPDK. Notably however they reported
7us when using TCP/IP from the FPGA rather than their own connection oriented
network protocol. On the other hand their peak replication throughput is 6M
replications/sec on 66 Gbps of aggregate network bandwith, comparable to the
same mark obtained by Cyclone on an aggregate 40 Gbps of network bandwidth when
not using the log on the SSD. We believe the general applicability afforded by
our software only solution together with the fact that we provide persistence by
design makes it a compelling alternative to an FPGA based solution even given
the larger latency.
